{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-07 Paper/main.tex
Changes from meetings

	\section{Theory}\label{Theory}
	In multivariate regression, data is often observed in the form of elements from Euclidean space, $\mathbb{R}^p$. However, the statistics derived from infinite-dimensional random functions cannot be defined on a finite-dimensional space. Therefore, it is necessary to introduce some concepts and extend known aspects of linear regression theory to include functional objects. One integral concept in inferential statistics is random variables. Paraphrasing a definition by \cite{bauer_wahrscheinlichkeitstheorie_2020}, a random variable $X:\Omega \rightarrow \Omega'$ is an $\mathcal{A} \text{-} \mathcal{A'} \text{-measurable}$ function, where $(\Omega, \mathcal{A}, P)$ is a probability space and $(\Omega', \mathcal{A'})$ is a measure space.\\
	A typical case known to every undergraduate student of economics is $(\Omega', \mathcal{A'}) = (\mathbb{R}, \mathcal{B})$, where $\mathcal{B}$ is the canonical $\sigma$-algebra on the real numbers. As a first intuition, it is possible to imagine a similar concept where a random variable does not realize as an element of the real numbers but as a function in a function space. A formalization of this idea makes some more theoretical considerations necessary. The following theoretical introduction closely follows chapters 2.3 and 2.4 from \cite{hsing_theoretical_2015}.

@@

	\paragraph{Monomial Basis}
	The Monomial basis are trivially decompositions of every polynomial function. {\color{red} Continue Work here!}	
	For functions that fall into the class of entire functions, we can express the function as a potentially infinite sum of weighted monomials. Let $f(t)$ be an entire function, e.g. the exponential function or a trigonometric function, then we can write the following.

@@

	Due to the chosen implementation for our simulation, we will limit our paper to the case of $\alpha = 0$. However, a different choice of $\alpha$ could lead to improvements in performance of the monomial basis in later parts of this paper.
	There exist collinearity problems of this basis system since the monomial basis functions become more correlated to each other as the degrees increase, which would result in numerically unstable situation. This restricts the number of basis functions. However, the lower order monomial function makes it impracticable to capture pronounced local peculiarities. It finally leads to an undesirable behavior at the tails such as Taylor expansion (cf. \cite{ramsay_functional_2005}). Therefore, regarding the use, this may be useful for relatively simple functions.

@@

	Its elements exhibit a cyclical behavior which is useful to expand functions that represent a periodic or seasonal underlying process. Additionally, the Fourier basis is especially suitable to expand functions with a similar curvature order across the domain, generally resulting in uniformly smooth expansions.

@@

	\subsection{Approximation and Smoothing via Basis Truncation}
	As mentioned above, the realized curves can be expressed in terms of a chosen functional basis. For this expansion, it is technically possible to use a complete basis of $\mathbb{L}^2[0,1]$. In many cases this is not a desirable approach as the expansion with too many basis functions introduces high amounts of variance in pronounced local variations of the curves or can even led to the approximation of noise in the sample curves, which possibly interrupts the analysis. {\color{red} Therefore, methods such as acceleration penalties have to be employed to smooth the function.} On the other hand, important information on the curves could be missed by using a number of basis functions that is too small. In that sense, this discussion is subject to the Bias-Variance tradeoff. It challenges the researcher to seek a point at which they truncate the basis function to remove noise and, at the same time, not to introduce too much bias by maintaining significant fluctuations of the curves. The basis expansion with truncation is defined by
	\begin{equation}
		X(\omega_0) = x(t) = \sum_{j \in \mathcal{I}} A_j(\omega_0) \phi_j(t) = \sum_{j = 1}^{L} A_j(\omega_0) \phi_j(t) + \delta(t) \approx \sum_{j = 1}^{L} A_j(\omega_0) \phi_j(t)
	\end{equation}
	where $\delta(t)$ is the truncation error and $L \leq \max\limits_{j \in \mathcal{I}}(j)$ for all $L \in \mathcal{I}$. The number $L$ can be chosen subjectively, but also through applying a data-driven method like Cross-Validation, which aims to minimize the Mean Squared Error (MSE) or other criteria. The influence of this truncation on the prediction accuracy of different estimates is the subject of the simulation study.

@@

	The parameter $M$ corresponds to the chosen number of principal components and therefore constitutes another choice in the approximation.

@@

	\subsection{Generating Similar Curves}
	To avoid small sample problems, we generated 200 similar curves from the gasoline dataset, motivated by \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. First, the initial curves are expressed in terms of a generated cubic B-spline basis which is created using 50 knots. In the R implementation of the fda package that is used, these 50 knots account for 52 basis functions (50+4-2). These smooth curves are then centered, before applying the \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. It is assumed that the scores follow a normal distribution, so the new realizations for the scores are drawn from a multivariate normal distribution $\tilde{\xi} = \left(\tilde{\xi}_{1},\: \dots \:, \tilde{\xi}_{M}\right)' \sim \mathcal{N}(0_M, \; diag(\hat{\lambda}_1,\: \dots\:, \hat{\lambda}_M))$. Finally, we obtain the generated curves $NIR_{sim}$

	\begin{equation}
		\tilde{X}(\omega)(t) = \hat{\mu}(t) + \sum_{m = 1}^{M} \tilde{\xi}_m(\omega) \tilde{\nu}^{m,L}(t)
	\end{equation}

	where $\tilde{X}(\omega)(t)$, $\hat{\mu}(t)$ and $\tilde{\nu}^{m,L}(t)$  are approximated as vectors in $\mathbb{R}^{401}$ for $L =$ 30 eigenfunctions.

@@

		 Two different error-terms $\epsilon$ were created by first generating an $i.i.d.$ standard normal error term and then multiplying it by two error variations $\sigma_e $. The error variations represent different signal-to-noise ratios to test the methods with low and high amounts of noise. They are created such that the squared multiple correlation coefficient $R^2 = var(Xf) / (var(Xf) + \sigma^2_{e})$ is equal to 0.9 and 0.6. The two error terms are then used to generate two sets of responses for $f \in \{f_1(t), f_2(t)\}$

@@

		\begin{equation}
			\begin{split}
				\hat{\xi}_{i}^{m, \mathcal{H}} &=  \int_{0}^{1} {\color{red} \left(X_{i}^{\mathcal{H}}(s) - \hat{\mu}^{\mathcal{T}}(s)\right)} {\color{blue}\hat{\nu}^{m, \mathcal{T}} }\mathrm{d}s) 
			    = \int_{0}^{1} {\color{red}\left(\sum_{j \in \mathcal{I}} a_{i,j}^{\mathcal{H}} \phi_j(s)\right)} {\color{blue}\left(\sum_{k \in \mathcal{L}} b_{k}^{m, \mathcal{T}} \psi_{k}(s)\right)} \mathrm{d}s \\
				 &= \sum_{j \in \mathcal{I}} \left[ a_{i,j}^{\mathcal{H}}\sum_{k \in \mathcal{L}}  b_{k}^{m, \mathcal{T}} \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s\right]
			\end{split}
		\end{equation}
		with truncation parameter $m$.
		\vspace{0.2cm}

@@

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{\textbf{Estimation using Basis-Representation}}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory.

@@

>>>>

	\section{Theory}\label{Theory}
	{\color{red} Introductory Sentence Here!} In multivariate regression, data is often observed in the form of elements from Euclidean space, $\mathbb{R}^p$. However, the statistics derived from infinite-dimensional random functions cannot be defined on a finite-dimensional space. Therefore, it is necessary to introduce some concepts and extend known aspects of linear regression theory to include functional objects. One integral concept in inferential statistics is random variables. Paraphrasing a definition by \cite{bauer_wahrscheinlichkeitstheorie_2020}, a random variable $X:\Omega \rightarrow \Omega'$ is an $\mathcal{A} \text{-} \mathcal{A'} \text{-measurable}$ function, where $(\Omega, \mathcal{A}, P)$ is a probability space and $(\Omega', \mathcal{A'})$ is a measure space.
	A typical case is $(\Omega', \mathcal{A'}) = (\mathbb{R}, \mathcal{B})$, where $\mathcal{B}$ is the canonical $\sigma$-algebra on the real numbers. As a first intuition, it is possible to imagine a similar concept where a random variable does not realize as an element of the real numbers but as a function in a function space. A formalization of this idea makes some more theoretical considerations necessary. The following theoretical introduction closely follows chapters 2.3 and 2.4 from \cite{hsing_theoretical_2015}.

@@

	\paragraph{Monomial Basis}
	For functions that fall into the class of entire functions, we can express the function as a potentially infinite sum of weighted monomials. Let $f(t)$ be an entire function, e.g. a polynomial, the exponential function or a trigonometric function, then we can write the following.

@@

	Due to the chosen implementation for our simulation, we will limit our paper to the case of $\alpha = 0$. However, a different choice of $\alpha$ could lead to improvements in performance of the monomial basis in later parts of this paper.
	Collinearity problems arise since the monomials become more correlated to each other as their degree increases, which can result in numerically unstable calculations. This restricts the number of basis functions that can be used in the estimation procedures limiting its ability to capture pronounced local peculiarities. It can also lead to undesirable behavior away from the point of evaluation of the monomials (cf. \cite{ramsay_functional_2005}).

@@

	Its elements exhibit a cyclical behavior which is useful to expand functions that represent a periodic or seasonal underlying process. Additionally, the Fourier basis is especially suitable to expand functions with a similar curvature order across the domain, generally resulting in uniformly smooth expansions. {\color{red} Citation}

@@

	\subsection{Approximation and Smoothing via Basis Truncation}
	As mentioned above, the realized curves can be expressed in terms of a chosen functional basis. For this expansion, it is technically possible to use a complete basis of $\mathbb{L}^2[0,1]$. In many cases this is not a desirable approach as the expansion with too many basis functions introduces high amounts of variance in pronounced local variations of the curves or can even led to the approximation of noise in the sample curves, which possibly interrupts the analysis. {\color{red} Therefore, methods such as acceleration penalties have to be employed to smooth the function.} On the other hand, important information on the curves could be missed by using a number of basis functions that is too small. In that sense, this discussion is subject to the Bias-Variance tradeoff. It challenges the researcher to seek a point at which they truncate the basis function to remove noise and, at the same time, not to introduce too much bias by maintaining significant fluctuations of the curves. The basis expansion with truncation is defined by
	\begin{equation}
		X(\omega_0) = x(t) = \sum_{j \in \mathcal{I}} A_j(\omega_0) \phi_j(t) = \sum_{j = 1}^{L} A_j(\omega_0) \phi_j(t) + \delta(t) \approx \sum_{j = 1}^{L} A_j(\omega_0) \phi_j(t)
	\end{equation}
	where $\delta(t)$ is the truncation error and $L \leq \max\limits_{j \in \mathcal{I}}(j)$ for all $L \in \mathcal{I}$. The number $L$ can be chosen subjectively, but also through applying a data-driven method like Cross-Validation, which aims to minimize the Mean Squared Error (MSE) or other criteria. The influence of this truncation on the prediction accuracy of different estimates is the subject of the simulation study. {\color{red} Add text for figure}

@@

	The parameter $M$ corresponds to the chosen number of principal components and therefore constitutes another choice in the approximation. {\color{red} Write equation with limited number of principal components. Truncation parameter M instead of N-1}

@@

	\subsection{Generating Similar Curves}
	To avoid small sample problems, we generated 200 similar curves from the gasoline dataset, motivated by \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. First, the initial curves are expressed in terms of a generated cubic B-spline basis which is created using 50 knots. In the R implementation of the fda package that is used, these 50 knots account for 52 basis functions (50+4-2). These smooth curves are then centered, before applying the \hyperlink{KL}{Karhunen-Lo\'{e}ve Expansion}. It is assumed that the scores follow a normal distribution, so the new realizations for the scores are drawn from a multivariate normal distribution $\mathring{\xi} = \left(\mathring{\xi}_{1},\: \dots \:, \mathring{\xi}_{M}\right)' \sim \mathcal{N}(0_M, \; diag(\hat{\lambda}_1,\: \dots\:, \hat{\lambda}_M))$. Finally, we obtain the generated curves $NIR_{sim}$

	\begin{equation}
		\mathring{X}(\omega)(t) = \hat{\mu}(t) + \sum_{m = 1}^{M} \mathring{\xi}^m(\omega) \tilde{\nu}^{m,L}(t)
	\end{equation}

	where $\mathring{X}(\omega)(t)$, $\hat{\mu}(t)$ and $\tilde{\nu}^{m,L}(t)$  are approximated as vectors in $\mathbb{R}^{401}$ for $M =$ 30 FPC's.

@@

		 Two different error-terms $\epsilon$ were created by first generating an $i.i.d.$ standard normal error term and then multiplying it by two error variations $\sigma_e $. The error variations represent different signal-to-noise ratios to test the methods with low and high amounts of noise. They are created such that the squared multiple correlation coefficient $R^2 = var(\langle X, f\rangle) / (var(\langle X, f\rangle) + \sigma^2_{e})$ is equal to 0.9 and 0.6. The two error terms are then used to generate two sets of responses for $f \in \{f_1(t), f_2(t)\}$

@@

		\begin{equation}
			\begin{split}
				\hat{\xi}_{i}^{m, \mathcal{H}} &\approx  \int_{0}^{1} {\color{red} \left(X_{i}^{\mathcal{H}}(s) - \hat{\mu}^{\mathcal{T}}(s)\right)} {\color{blue}\hat{\nu}^{m, \mathcal{T}} }\mathrm{d}s) 
			    = \int_{0}^{1} {\color{red}\left(\sum_{j = 1}^{L} a_{i,j}^{\mathcal{H}} \phi_j(s)\right)} {\color{blue}\left(\sum_{k = 1}^{L} b_{k}^{m, \mathcal{T}} \phi_{k}(s)\right)} \mathrm{d}s \\
				 &\approx \sum_{j = 1}^{L} \left[ a_{i,j}^{\mathcal{H}}\sum_{k = 1}^{L}  b_{k}^{m, \mathcal{T}} \int_{0}^{1} \phi_j(s) \psi_{k}(s)\mathrm{d}s\right]
			\end{split}
		\end{equation}
		with truncation parameter $m$.
		\vspace{0.2cm}

@@

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}(t)$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{\textbf{Estimation using Basis-Representation}}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory.

@@

	\paragraph{Larger Range for the Number of Basis Functions for FPCR}

	\paragraph{Larger Range for the Number of Principal Components}

	\paragraph{Input from Physics Analysis}
